---
title: "R_Assignment"
author: "Elizabeth Glynne"
date: "10/16/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
_RMD for R Assignment: This is the markdown file for my assignment. It will describe my general workflow for this project._

### Link R and Github

I created a repository on Github and then created a new project from Version control.

## Part I: Data inspection
### Step 1: Import Data

I copied the data from the UNIX assignment into my R_Assignment folder using my local repositories. I then pushed them to my global repository and then pulled the repository through my Rstudio interface.

Loading in the genotypes data:
```{r message=FALSE, warning=FALSE, results=FALSE}
library(tidyverse)
fang <- read_tsv(file = "fang_et_al_genotypes.txt", col_names = TRUE)
```

Loading in the SNPs data:
```{r warning=FALSE, message=FALSE, results=FALSE}
snp <- read_tsv(file = "snp_position.txt")
```

### Step 2: Inspect Genotypes
View the data file:

```{r, message=FALSE, warning=FALSE, results=FALSE}
View(fang)
```
In viewing the data file, we can see that it is a table with headers. 

This file is currently a list:
```{r, message=FALSE, warning=FALSE, results=FALSE}
typeof(fang)
```
This file has 986 variables, with 2782 rows and 986 columns.
```{r, message=FALSE, warning=FALSE, results=FALSE}
length(fang)
dim(fang)
```
By first storing the attributes of the genotype file as an object, I can further analyze each attribute of the list.  
```{r, results=FALSE}
genotype_attributes <- attributes(fang)
length(genotype_attributes$names)
length(genotype_attributes$class)
length(genotype_attributes$row.names)
typeof(genotype_attributes$names)
typeof(genotype_attributes$row.names)
typeof(genotype_attributes$class)
genotype_attributes$class
```
We can see that the lengths of each attribute, how the names correspond to the number of columns (986) and the row names correspond to the number of rows (2782). However, the length of the classes (4) corresponds to the various components being a special table dataframe, table dataframe, table, and a data frame.


### Step 3: Inspect SNPs
View the data file:

```{r, message=FALSE, warning=FALSE, results=FALSE}
View(snp)
```
In viewing the data file, we can see that it is a table with headers. 

This file is currently a list:
```{r}
typeof(snp)
```
This file has 15 variables, with 983 rows and 15 columns of data.
```{r}
length(snp)
dim(snp)
```
By first storing the attributes of the snps file as an object, I can further analyze each attribute of the list.  
```{r results=FALSE}
snp_attributes <- attributes(snp)
length(snp_attributes$names)
length(snp_attributes$class)
length(snp_attributes$row.names)
typeof(snp_attributes$names)
typeof(snp_attributes$row.names)
typeof(snp_attributes$class)
snp_attributes$class
```
We can see that the lengths of each attribute, how the names correspond to the number of columns (15) and the row names correspond to the number of rows (983). However, the length of the classes (4) corresponds to the various components being a special table dataframe, table dataframe, table, and a data frame.


## Part II: Data Processing
### Step 1: Prep files
Sort the SNPs by ID
```{r}
library(tidyr)
sorted_snp<-arrange(snp, SNP_ID)
```

### Step 1: Define groups
Pull out Maize (ZMMIL, ZMMLR, and ZMMMR)
```{r}
maize<- filter(fang, Group %in% c("ZMMIL", "ZMMLR", "ZMMMR"))
```

Pull out Teosinte (ZMPBA, ZMPIL, ZMPJA)

```{r}
teosinte<- filter(fang, Group %in% c("ZMPBA", "ZMPIL", "ZMPJA"))
```

### Step 2: Transpose file Maize and Teosinte
```{r}
Tmaize<- as.data.frame(t(maize))
Tteosinte <-as.data.frame(t(teosinte))
```

### Step 3: Join SNP and Genotype files for both Maize and Teosinte

Join together the files into a new table, with the first three columns from the sorted_snp file (SNP_ID, Chromosome, and Position). This will be used with the creation of the next two files. 

```{r}
join_start<-bind_cols(data.frame(SNP_ID = sorted_snp$SNP_ID),data.frame(Chromosome = sorted_snp$Chromosome),data.frame(Position = sorted_snp$Position))
```
### Step 4: Create the 10 chromosome files for Maize + 1 for unknown and 1 for multiple

Change dataframe so the row names are now in a column called SNP_ID. 
Join together the Tmaize file with the join_start file, resulting in the join_maize file which will be used for later. 

```{r, include=FALSE, message=FALSE}
Tmaize<-tibble::rownames_to_column(Tmaize, "SNP_ID") 
join_maize <- full_join(join_start, Tmaize, by= "SNP_ID")

```


Create join_maize1 file, changing "?/?" to "?". Also, create join_maize2 file, changing "?" to '-".
```{r}
join_maize <- data.frame(lapply(join_maize, as.character),
                            stringsAsFactors=FALSE)
join_maize1 <- data.frame(sapply(join_maize,function(x) 
                  {x <- gsub("?/?","?",x,fixed=TRUE)}))
join_maize2 <- data.frame(sapply(join_maize1,function(x) 
                  {x <- gsub("?","-",x,fixed=TRUE)}))

```

#### Part i: Create file with unknown data
Create and write a file with all SNPs with unknown positions in the genome:

```{r}
maize_unknown <- filter(join_maize, Chromosome == "unknown")
write.table(maize_unknown,"m_Chrom_Unknown", sep="\t", row.names=FALSE)
```

#### Part ii: Create file with multiple data
Create and write a file with all SNPs with multiple positions in the genome:
```{r}
maize_multiple <- filter(join_maize, Chromosome == "multiple")
write.table(maize_multiple,"m_Chrom_Multiple", sep="\t", row.names=FALSE)
```

#### Part iii: Create a file for each chromosome with positions in increasing order and missing data denoted as "?"

Sort the join_maize file into 10 new data frames, one for each chromosome:
```{r, results=FALSE}
df.names <- paste("maize_chrom", 1:10,sep="_")
for (i in 1:10) {
  d.frame <- filter(join_maize1, Chromosome == i)
  assign(df.names[i],d.frame)
}
```

Order the files based on the increasing postion values:
```{r, messages=FALSE}
maize_chrom_1_s<- maize_chrom_1[order(as.numeric(as.character(maize_chrom_1$Position))),]
maize_chrom_2_s<- maize_chrom_2[order(as.numeric(as.character(maize_chrom_2$Position))),]
maize_chrom_3_s<- maize_chrom_3[order(as.numeric(as.character(maize_chrom_3$Position))),]
maize_chrom_4_s<- maize_chrom_4[order(as.numeric(as.character(maize_chrom_4$Position))),]
maize_chrom_5_s<- maize_chrom_5[order(as.numeric(as.character(maize_chrom_5$Position))),]
maize_chrom_6_s<- maize_chrom_6[order(as.numeric(as.character(maize_chrom_6$Position))),]
maize_chrom_7_s<- maize_chrom_7[order(as.numeric(as.character(maize_chrom_7$Position))),]
maize_chrom_8_s<- maize_chrom_8[order(as.numeric(as.character(maize_chrom_8$Position))),]
maize_chrom_9_s<- maize_chrom_9[order(as.numeric(as.character(maize_chrom_9$Position))),]
maize_chrom_10_s<- maize_chrom_10[order(as.numeric(as.character(maize_chrom_10$Position))),]

```

Write files:
```{r}
write.table(maize_chrom_1_s,"m_Chrom1", sep="\t", row.names=FALSE)
write.table(maize_chrom_2_s,"m_Chrom2", sep="\t", row.names=FALSE)
write.table(maize_chrom_3_s,"m_Chrom3", sep="\t", row.names=FALSE)
write.table(maize_chrom_4_s,"m_Chrom4", sep="\t", row.names=FALSE)
write.table(maize_chrom_5_s,"m_Chrom5", sep="\t", row.names=FALSE)
write.table(maize_chrom_6_s,"m_Chrom6", sep="\t", row.names=FALSE)
write.table(maize_chrom_7_s,"m_Chrom7", sep="\t", row.names=FALSE)
write.table(maize_chrom_8_s,"m_Chrom8", sep="\t", row.names=FALSE)
write.table(maize_chrom_9_s,"m_Chrom9", sep="\t", row.names=FALSE)
write.table(maize_chrom_10_s,"m_Chrom10", sep="\t", row.names=FALSE)

```

#### Part iv: Create a file for each chromosome with positions in decreasing order and missing data denoted as "-"
Sort the join_maize file into 10 new data frames, one for each chromosome. 

```{r}
df.names <- paste("maize_chrom_de", 1:10,sep="_")
for (i in 1:10) {
  d.frame <- filter(join_maize2, Chromosome == i)
  assign(df.names[i],d.frame)
}
```

Order the files based on the decreasing position values:

```{r}
maize_chrom_de_1<- maize_chrom_de_1[order(as.numeric(as.character(maize_chrom_de_1$Position)), decreasing=TRUE),]

maize_chrom_de_2<- maize_chrom_de_2[order(as.numeric(as.character(maize_chrom_de_2$Position)), decreasing=TRUE),]

maize_chrom_de_3<- maize_chrom_de_3[order(as.numeric(as.character(maize_chrom_de_3$Position)), decreasing=TRUE),]

maize_chrom_de_4<- maize_chrom_de_4[order(as.numeric(as.character(maize_chrom_de_4$Position)), decreasing=TRUE),]

maize_chrom_de_5<- maize_chrom_de_5[order(as.numeric(as.character(maize_chrom_de_5$Position)), decreasing=TRUE),]

maize_chrom_de_6<- maize_chrom_de_6[order(as.numeric(as.character(maize_chrom_de_6$Position)), decreasing=TRUE),]

maize_chrom_de_7<- maize_chrom_de_7[order(as.numeric(as.character(maize_chrom_de_7$Position)), decreasing=TRUE),]

maize_chrom_de_8<- maize_chrom_de_8[order(as.numeric(as.character(maize_chrom_de_8$Position)), decreasing=TRUE),]

maize_chrom_de_9<- maize_chrom_de_9[order(as.numeric(as.character(maize_chrom_de_9$Position)), decreasing=TRUE),]

maize_chrom_de_10<- maize_chrom_de_10[order(as.numeric(as.character(maize_chrom_de_10$Position)), decreasing=TRUE),]

```

Write files
```{r}
write.table(maize_chrom_de_1,"m_Chrom1_de", sep="\t", row.names=FALSE)
write.table(maize_chrom_de_2,"m_Chrom2_de", sep="\t", row.names=FALSE)
write.table(maize_chrom_de_3,"m_Chrom3_de", sep="\t", row.names=FALSE)
write.table(maize_chrom_de_4,"m_Chrom4_de", sep="\t", row.names=FALSE)
write.table(maize_chrom_de_5,"m_Chrom5_de", sep="\t", row.names=FALSE)
write.table(maize_chrom_de_6,"m_Chrom6_de", sep="\t", row.names=FALSE)
write.table(maize_chrom_de_7,"m_Chrom7_de", sep="\t", row.names=FALSE)
write.table(maize_chrom_de_8,"m_Chrom8_de", sep="\t", row.names=FALSE)
write.table(maize_chrom_de_9,"m_Chrom9_de", sep="\t", row.names=FALSE)
write.table(maize_chrom_de_10,"m_Chrom10_de", sep="\t", row.names=FALSE)

```


### Step 5: Create the 10 chromosome files for Teosinte + 1 for unknown and 1 for multiple
Change dataframe so the row names are now in a column called SNP_ID.
Join together the Tteosinte file with the join_start file, resulting in
the join_teosinte file which will be used for later analysis. 

```{r}
Tteosinte<-tibble::rownames_to_column(Tteosinte, "SNP_ID") 
join_teosinte <- full_join(join_start, Tteosinte, by= "SNP_ID")
```


Create join_teosinte1 file, changing "?/?" to "?". Also, create join_teosinte2 file, changing "?" to '-".
```{r}
join_teosinte <- data.frame(lapply(join_teosinte, as.character),
                            stringsAsFactors=FALSE)
join_teosinte1 <- data.frame(sapply(join_teosinte,function(x) 
                  {x <- gsub("?/?","?",x,fixed=TRUE)}))
join_teosinte2 <- data.frame(sapply(join_teosinte1,function(x) 
                  {x <- gsub("?","-",x,fixed=TRUE)}))

```

#### Part i: Create file with unknown data
Create and write a file with all SNPs with unknown positions in the genome:

```{r}
teosinte_unknown <- filter(join_teosinte, Chromosome == "unknown")
write.table(teosinte_unknown,"t_Chrom_Unknown", sep="\t", row.names=FALSE)
```

#### Part ii: Create file with multiple data
Create and write a file with all SNPs with multiple positions in the genome:
```{r}
teosinte_multiple <- filter(join_teosinte, Chromosome == "multiple")
write.table(teosinte_multiple,"m_Chrom_Multiple", sep="\t", row.names=FALSE)
```

#### Part iii: Create a file for each chromosome with positions in increasing order and missing data denoted as "?"

Sort the join_teosinte file into 10 new data frames, one for each chromosome:
```{r, results=FALSE}
df.names <- paste("teosinte_chrom", 1:10,sep="_")
for (i in 1:10) {
  d.frame <- filter(join_teosinte1, Chromosome == i)
  assign(df.names[i],d.frame)
}
```

Order the files based on the increasing postion values:
```{r, messages=FALSE}
teosinte_chrom_1_s<-teosinte_chrom_1[order(as.numeric(as.character(teosinte_chrom_1$Position))),]

teosinte_chrom_2_s<-teosinte_chrom_2[order(as.numeric(as.character(teosinte_chrom_2$Position))),]

teosinte_chrom_3_s<-teosinte_chrom_3[order(as.numeric(as.character(teosinte_chrom_3$Position))),]

teosinte_chrom_4_s<-teosinte_chrom_4[order(as.numeric(as.character(teosinte_chrom_4$Position))),]

teosinte_chrom_5_s<-teosinte_chrom_5[order(as.numeric(as.character(teosinte_chrom_5$Position))),]

teosinte_chrom_6_s<-teosinte_chrom_6[order(as.numeric(as.character(teosinte_chrom_6$Position))),]

teosinte_chrom_7_s<-teosinte_chrom_7[order(as.numeric(as.character(teosinte_chrom_7$Position))),]

teosinte_chrom_8_s<-teosinte_chrom_8[order(as.numeric(as.character(teosinte_chrom_8$Position))),]

teosinte_chrom_9_s<-teosinte_chrom_9[order(as.numeric(as.character(teosinte_chrom_9$Position))),]

teosinte_chrom_10_s<-teosinte_chrom_10[order(as.numeric(as.character(teosinte_chrom_10$Position))),]


```

Write files:
```{r}
write.table(teosinte_chrom_1_s,"t_Chrom1", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_2_s,"t_Chrom2", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_3_s,"t_Chrom3", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_4_s,"t_Chrom4", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_5_s,"t_Chrom5", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_6_s,"t_Chrom6", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_7_s,"t_Chrom7", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_8_s,"t_Chrom8", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_9_s,"t_Chrom9", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_10_s,"t_Chrom10", sep="\t", row.names=FALSE)

```

#### Part iv: Create a file for each chromosome with positions in decreasing order and missing data denoted as "-"
Sort the join_maize file into 10 new data frames, one for each chromosome. 

```{r}
df.names <- paste("teosinte_chrom_de", 1:10,sep="_")
for (i in 1:10) {
  d.frame <- filter(join_teosinte2, Chromosome == i)
  assign(df.names[i],d.frame)
}
```

Order the files based on the decreasing position values:

```{r}
teosinte_chrom_de_1<- teosinte_chrom_de_1[order(as.numeric(as.character(teosinte_chrom_de_1$Position)), decreasing=TRUE),]

teosinte_chrom_de_2<- teosinte_chrom_de_2[order(as.numeric(as.character(teosinte_chrom_de_2$Position)), decreasing=TRUE),]

teosinte_chrom_de_3<- teosinte_chrom_de_3[order(as.numeric(as.character(teosinte_chrom_de_3$Position)), decreasing=TRUE),]

teosinte_chrom_de_4<- teosinte_chrom_de_4[order(as.numeric(as.character(teosinte_chrom_de_4$Position)), decreasing=TRUE),]

teosinte_chrom_de_5<- teosinte_chrom_de_5[order(as.numeric(as.character(teosinte_chrom_de_5$Position)), decreasing=TRUE),]

teosinte_chrom_de_6<- teosinte_chrom_de_6[order(as.numeric(as.character(teosinte_chrom_de_6$Position)), decreasing=TRUE),]

teosinte_chrom_de_7<- teosinte_chrom_de_7[order(as.numeric(as.character(teosinte_chrom_de_7$Position)), decreasing=TRUE),]

teosinte_chrom_de_8<- teosinte_chrom_de_8[order(as.numeric(as.character(teosinte_chrom_de_8$Position)), decreasing=TRUE),]

teosinte_chrom_de_9<- teosinte_chrom_de_9[order(as.numeric(as.character(teosinte_chrom_de_9$Position)), decreasing=TRUE),]

teosinte_chrom_de_10<- teosinte_chrom_de_10[order(as.numeric(as.character(teosinte_chrom_de_10$Position)), decreasing=TRUE),]

```

Write files
```{r}
write.table(teosinte_chrom_de_1,"t_Chrom1_de", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_de_2,"t_Chrom2_de", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_de_3,"t_Chrom3_de", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_de_4,"t_Chrom4_de", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_de_5,"t_Chrom5_de", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_de_6,"t_Chrom6_de", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_de_7,"t_Chrom7_de", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_de_8,"t_Chrom8_de", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_de_9,"t_Chrom9_de", sep="\t", row.names=FALSE)
write.table(teosinte_chrom_de_10,"t_Chrom10_de", sep="\t", row.names=FALSE)

```


## Part III: Data Visualization

### SNPs per chromosom


### Missing data and amount of heterozygosity



### "Your own visualization
